---
title: "Report6"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Load the R packages

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)
library(readr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(tibble)
library(readxl)
```

## Preparations: read the Excel data and do some quick renaming and check what happened at the 1956 games:

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Read data
athletes <- read_excel("olympics.xlsx", sheet="athletes")
games    <- read_excel("olympics.xlsx", sheet="games")
country  <- read_excel("olympics.xlsx", sheet="country")
medals   <- read_excel("olympics.xlsx", sheet="medals")
# Rename column for later JOINS
athletes <- athletes %>%
  rename(athlete_id = ID)
# What happened at the 1956 games?
games %>% filter(Year==1956)
```
There were actually two Summer Games in 1956 - one in Melbourne and one in Stockholm!

## Part 1

Have some athletes competed for different countries over time?
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Do the necessary JOINS
part1 <- athletes %>%
  full_join(country)
# Check summary statistics and remove duplicates
part1 <- part1 %>%
  group_by(Name, NOC) %>%
  select(Name, NOC) %>%
  distinct() %>%
  group_by(Name) %>%
  summarize(NUMBER_OF_COUNTRIES = n())
```
Indeed, `r part1 %>% filter(NUMBER_OF_COUNTRIES>1) %>% nrow()` athletes had various country affiliations.

## Part 2

Who are the ten athletes that took part in most games?
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Do the necessary JOINS
part2 <- athletes %>%
  full_join(country)
# Summary statistics
part2 %>%
  group_by(Name) %>%
  summarize(NUMBER_OF_GAMES = n()) %>%
  arrange(desc(NUMBER_OF_GAMES)) %>%
  head(10)
```
## Part 3

What athlete(s) kept a Gold medal for the longest time?
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Do the necessary JOINS
part3 <- athletes %>%
  full_join(medals) %>%
  filter(Medal == 'Gold') %>%
  group_by(Name, Event) %>%
  mutate(NUMBER_OF_GOLD_BY_EVENT = n()) %>%
  ungroup() %>% 
  filter(NUMBER_OF_GOLD_BY_EVENT>1) %>%
  arrange(desc(NUMBER_OF_GOLD_BY_EVENT)) %>%
  select(Name, Sex, Games, Team, Sport, Medal, NUMBER_OF_GOLD_BY_EVENT)
  head(10)
# Print
part3
```
Considering Bobb'y comment "With keeping a gold medal we mean a gold medal on the same event", the tibble above shows that the Hungarian athlete *Aladr Gerevich (-Gerei)* won 6 gold medals in *Fencing Men's Sabre, Team* during at the following games: `r part3 %>% filter(Name=='Aladr Gerevich (-Gerei)') %>% pull(Games)`.

## Part 4

Based on the tibble of Part 3, Hungary kept a Gold medal for the longest time.

## Part 5

Who are the ten athletes that competed in the most events (some athletes take part in more than one event during games) ?

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Do the necessary JOINS
part5 <- athletes %>%
  full_join(country) %>% 
  full_join(medals) %>%
  distinct() %>%
  group_by(Name, Games) %>%
  summarise(NUMBER_OF_COMPETITIONS = n()) %>%
  arrange(desc(NUMBER_OF_COMPETITIONS)) %>%
  head(10)
# Print
part5
```

## Part 6

Create a new table showing the number of medals per country (rows) and per year (column). Keep only the 15 countries with the most medals overall.

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Compute the 15 countries with most medals
part6_15countries <- games %>%
  full_join(medals) %>% 
  full_join(athletes) %>% 
  select(Year, Team, Medal) %>%
  filter((Medal=='Gold') | (Medal=='Silver') | (Medal=='Bronze')) %>% 
  group_by(Team) %>%
  summarize(TOTAL_MEDALS_OVERALL = n()) %>%
  arrange(desc(TOTAL_MEDALS_OVERALL)) %>%
  head(15)
# Print
part6_15countries
```
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Pull the country names for later comparison
top_countries <- pull(part6_15countries, Team)
# Now the same joins again but keep only the 15 countries with most medals
part6_medals <- games %>%
  full_join(medals) %>% 
  full_join(athletes) %>% 
  select(Year, Team, Medal) %>%
  filter(str_detect(Team, paste(top_countries, collapse = "|"))) %>%
  filter(str_detect(Team, "-", negate = TRUE)) %>%
  filter((Medal=='Gold') | (Medal=='Silver') | (Medal=='Bronze')) %>%
  group_by(Year, Team) %>%
  mutate(TOTAL_MEDALS = n()) %>%
  distinct() %>%
  ungroup() %>%
  arrange(Year) %>%
  select(Year, Team, TOTAL_MEDALS)
# Print a snippet
part6_medals %>% head(10)
```

## Part 7

Create a scatterplot showing the average height and weight of competitors per sport (one dot per sport).

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Do the JOINS and compute the quantities
part7 <- athletes %>%
  full_join(medals) %>%
  group_by(Sport) %>%
  mutate(AVERAGE_HEIGHT = mean(Height, na.rm = TRUE)) %>% 
  mutate(AVERAGE_WEIGHT = mean(Weight, na.rm = TRUE)) %>% 
  mutate(AVERAGE_BMI    = mean((Weight / ((Height/100)*(Height/100))), na.rm = TRUE)) %>% 
  select(AVERAGE_HEIGHT, AVERAGE_WEIGHT, AVERAGE_BMI, Sport) %>%
  ungroup() %>%
  distinct()
# Print some values
part7 %>% head(10)
```
Computer other variables
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Include the other quantities (nested, not piped)
MAX_AVERAGE_HEIGHT = pull(arrange(part7,desc(AVERAGE_HEIGHT))["Sport"])[1]
MIN_AVERAGE_HEIGHT = pull(arrange(part7,(AVERAGE_HEIGHT))["Sport"])[1]
MAX_AVERAGE_WEIGHT = pull(arrange(part7,desc(AVERAGE_WEIGHT))["Sport"])[1]
MIN_AVERAGE_WEIGHT = pull(arrange(part7,(AVERAGE_WEIGHT))["Sport"])[1]
MAX_AVERAGE_BMI    = pull(arrange(part7,desc(AVERAGE_BMI))["Sport"])[1]
MIN_AVERAGE_BMI    = pull(arrange(part7,(AVERAGE_BMI))["Sport"])[1]
```
The sport with the largest average height is `r MAX_AVERAGE_HEIGHT`.
The sport with the smaller average height is `r MIN_AVERAGE_HEIGHT`.
The sport with the largest average weight is `r MAX_AVERAGE_WEIGHT`.
The sport with the smaller average weight is `r MIN_AVERAGE_WEIGHT`.
The sport with the largest average BMI is `r MAX_AVERAGE_BMI`.
The sport with the smaller average BMI is `r MIN_AVERAGE_BMI`.

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.width=7.5, fig.height=14}
part7_filter <- filter(part7, Sport==MAX_AVERAGE_HEIGHT)

ggplot(part7, aes(x=AVERAGE_HEIGHT, y=AVERAGE_WEIGHT, color=Sport)) +
  geom_point() +
  theme(legend.position="bottom") +
  geom_text(data = filter(part7, Sport==MAX_AVERAGE_HEIGHT), aes(label = Sport)) + 
  geom_text(data = filter(part7, Sport==MIN_AVERAGE_HEIGHT), aes(label = Sport)) + 
  geom_text(data = filter(part7, Sport==MAX_AVERAGE_WEIGHT), aes(label = Sport)) + 
  geom_text(data = filter(part7, Sport==MIN_AVERAGE_WEIGHT), aes(label = Sport)) + 
  geom_text(data = filter(part7, Sport==MAX_AVERAGE_BMI), aes(label = Sport)) + 
  geom_text(data = filter(part7, Sport==MIN_AVERAGE_BMI), aes(label = Sport))
```
## Part 8

Create a line plot showing the number of medals given by year (one line for Gold, one line for Silver and one line for Bronze). Does it change over time? 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Do the JOINS
part8 <- games %>%
  full_join(medals) %>%
  group_by(Year, Medal) %>%
  mutate(NUMBER_OF_MEDALS = n()) %>%
  ungroup() %>%
  distinct()
```
Do the plot:
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Do the JOINS
part8 %>%
  filter((Medal=='Gold') | (Medal=='Silver') | (Medal=='Bronze')) %>%
  ggplot(aes(x=Year, y=NUMBER_OF_MEDALS, colour=Medal)) + 
  geom_line() +
  facet_grid(rows = vars(Season))
```
There is actually quite a significant change in the number of medals awarded over the years and in terms of Season (Summer vs. Winter).
 
