---
title: "Report2"
output: pdf_document
---

## Load the R packages

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)
library(readr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
```

## Load the data

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- read_csv("survey.csv")

data_types <- data %>% 
    dplyr::summarise_all(class) %>% 
    tidyr::gather(variable, class)

participants_per_university <- data %>%
  group_by(University) %>%
  summarize(total_participants = n())

```

## Part 1

They were **`r nrow(data)`** survey participants.

They survey is made up of the following data types: **`r unique(data_types$class)`**.

The top 3 universities with the most survey participants were
```{r echo=FALSE, results='asis'}
kable(head(participants_per_university,3))
```

## Part 2

A horizontal bar graph of the number of survey participants for the different positions given in the column *Position*:

```{r echo=TRUE, message=FALSE, warning=FALSE}
participants_per_position <- data %>%
  group_by(Position) %>%
  summarize(total_participants = n())

ggplot(participants_per_position, aes(x=Position, y=total_participants)) + geom_bar(position="dodge",stat="identity") + coord_flip()
```

## Part 3

```{r echo=TRUE, message=FALSE, warning=FALSE}
# define labels
label_s  <- "Small (0 to 10)"
label_m  <- "Medium (11 to 20)"
label_l  <- "Large (21 to 100)"
label_xl <- "Extra large (over 100)"

# Recreate the Sizes column
for (ind in seq(1,nrow(data))) {
  if (data$'Number of employees'[ind] <= 10) {
    data$'Sizes'[ind] <- label_s
  } else if (data$'Number of employees'[ind] <= 20) {
    data$'Sizes'[ind] <- label_m
  } else if (data$'Number of employees'[ind] <= 100) {
    data$'Sizes'[ind] <- label_l
  } else
    data$'Sizes'[ind] <- label_xl
}

participants_per_sizes <- data %>%
  group_by(Sizes) %>%
  summarize(total_participants = n())

# Plot
ggplot(participants_per_sizes, aes(x=Sizes, y=total_participants)) + geom_bar(position="dodge",stat="identity") + coord_flip()
```

## Part 4

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Mutate data with new aggregated values
data_share <- data %>%
  rename(
    n_total    = 'Number of employees',
    n_phds     = 'Number of PhD students',
    n_postdocs = 'Number of postdocs',
    n_profs    = 'Number of professors'
    ) %>%
    mutate(
      share_phds     = 100*n_phds/n_total,
      share_postdocs = 100*n_postdocs/n_total,
      share_profs    = 100*n_profs/n_total,
      share_total    = 100
      ) %>%
      select(X1,share_phds, share_postdocs, share_profs) %>%
      as_tibble()

data_gathered <- data_share %>%
  gather(key = "Type",
         value = "Percentage",
         share_phds,
         share_postdocs,
         share_profs)

data_gathered <- data_gathered %>%
  rename(Research_Group_ID = 'X1')

ggplot(data=data_gathered, aes(x=Research_Group_ID, y = Percentage, fill=Type)) + 
geom_bar(stat="identity") + coord_flip()
