---
title: "Report5"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Load the R packages

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)
library(readr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(httr)
library(tibble)
```

## Part 1

Sending some example queries to Movie DB with my API key:

```{r echo=TRUE, message=FALSE, warning=FALSE}
# What are the highest grossing dramas from 2010?
rm(data, data_content)
data <-  httr::GET("https://api.themoviedb.org/3/discover/movie?%with_genres=18&primary_release_year=2010&sort_by=revenue.desc&api_key=15079d94fcca66e5a038ec57e3f32000")
data_content <- httr::content(data)
```
The 3 highest grossing dramas from 2010 are `r purrr::pluck(data_content, "results", 1, "original_title")`, `r purrr::pluck(data_content, "results", 2, "original_title")`, `r purrr::pluck(data_content, "results", 3, "original_title")`.

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Have Will Ferrell and Liam Neeson even been in a movie together?
rm(data, data_content)
data <-  httr::GET("https://api.themoviedb.org/3/discover/movie?&with_cast=3896&with_cast=23659&api_key=15079d94fcca66e5a038ec57e3f32000")
data_content <- httr::content(data)
```
After inspecting the content of the GET response, the two actors Will Ferrell and Liam Neeson have not been in a movie together.
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Can you find kids movies with Tom Cruise in it?
rm(data, data_content)
data <-  httr::GET("https://api.themoviedb.org/3/discover/movie?&with_cast=500&certification_country=US&certification.lte=G&api_key=15079d94fcca66e5a038ec57e3f32000")
data_content <- httr::content(data)
```
After inspecting the content of the GET response, the Tom Cruise has not featured in a kids movie.

## Part 2

Search for the ID of "Pixar":
```{r echo=TRUE, message=FALSE, warning=FALSE}
rm(data, data_content)
data <-  httr::GET("https://api.themoviedb.org/3/search/company?api_key=15079d94fcca66e5a038ec57e3f32000&query=Pixar")
data_content <- httr::content(data, as = "parsed")
```
The Company ID is `r purrr::pluck(data_content, "results", 1, "id")`

## Part 3

Here we write a query that will search all the Pixar movies by descending revenue, which are then put into a tibble:
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Get the Pixar movies and sort them by descending revenue
rm(data, data_content)
data <-  httr::GET("https://api.themoviedb.org/3/discover/movie?%with_companies=3&sort_by=revenue.desc&api_key=15079d94fcca66e5a038ec57e3f32000")
data_content <- httr::content(data, as = "parsed")
# Convert them to a tibble
n_movies <- 20
titles <- vector("character", length(n_movies))
popularity <- vector("character", length(n_movies))
for (i in 1:n_movies) {
  titles[i] <- purrr::pluck(data_content, "results", i, "original_title")
  popularity[i] <- purrr::pluck(data_content, "results", i, "popularity")
}
movie_tibble <- as_tibble(titles) %>%
  rename(TITLE = value) %>%
  mutate(POPULARITY = popularity)
movie_tibble
```
The extraction of the movie title and the popularity could have been done with a combination of the map() and pluck() command, but for I opted for the sort of archaic solution with a simple for-loop.
